---
title: "Matematikresurser - en enkätundersökning vid Lunds universitet"
author: "Linda Hartman och Sara Maad Sasane"
format: 
  html:
    embed-resources: true
echo: false
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(skimr)
library(scales)

```

## Data

Data from 585 respondents to a survey. Data was exported to SPSS and cleaned in script `dm.R`. 

Free text answers are not included here.


```{r}
#| label: data

enkat <- readRDS('Data/enkat.rds') |> select(-starts_with("VAR"))
item_labels <- readRDS('Data/item_labels.rds')

```

```{r}
#| label: skim
#enkat |>  skim()
```

## Total time spent on this course


For all 
```{r}

nlbl= paste0(" n = ", enkat |>count()) 
enkat |>
  count(Tid_totalt_timmar) |>
  mutate(pct = n / sum(n), n_total = sum(n)) |>
  ungroup() |>
#  mutate(nlbl = paste0(" (n = ", n_total, ")")) |>
  ggplot(aes(x = Tid_totalt_timmar, y = pct, fill = Tid_totalt_timmar)) +
  geom_col() +
  geom_text(aes(label = n), vjust = -0.4, size = 3) +
  scale_y_continuous(labels = scales::percent, expand = expansion(mult = c(0, .12))) +
  scale_fill_brewer(palette = "Greens", direction = 1) +
  labs(x = NULL, y = NULL,  title = "Total hours spent",
       caption = str_wrap("Approximately, how many hours have you spent on your
                          mathematics studies in this course totally the last week? 
                          The question refers to the total time spent, 
                          consisting of scheduled time as well as 
                          self study.",width=120),
       subtitle = nlbl)  +
  theme_minimal() +
  theme(legend.position = "none")
```
## Total time: Separate for 7.5 hp courses (50% pace) B2, Matstat etc

```{r}
nlbl= paste0(" n = ", 
             enkat  |> filter(!is.na(course_gr)) |>count()) 

enkat |>
  filter(!is.na(course_gr)) |>
#  mutate(subject = factor(subject, levels = c("Matte LTH", "Matte NF", 
#  "Matstat LTH", "Matstat NF"))) |>
  count(course_gr, Tid_totalt_timmar) |>
  group_by(course_gr) |>
  mutate(pct = n / sum(n), n_total = sum(n)) |>
  ungroup() |>
  mutate(course_gr_lbl = fct_inorder(paste0(course_gr, " (n = ", n_total, ")"))) |>
  ggplot(aes(x = Tid_totalt_timmar, y = pct, fill = Tid_totalt_timmar)) +
  geom_col() +
  geom_text(aes(label = n), vjust = -0.4, size = 3) +
  scale_y_continuous(labels = scales::percent, 
                     expand = expansion(mult = c(0, .12))) +
  scale_fill_brewer(palette = "Greens", direction = 1) +
  facet_wrap(~ course_gr_lbl, nrow = 2) +
  labs(x = NULL, y = NULL, title = "Total study hours in this course",
       caption = str_wrap("Approximately, how many hours have you spent on your
                          mathematics studies in this course totally the last week? 
                          The question refers to the total time spent, 
                          consisting of scheduled time as well as 
                          self study.",width=120),
       ) +
    theme_minimal(base_size=14) +
  theme(legend.position = "none")
ggsave("Figures/total_time_sel.emf",width=12,height=6)
ggsave("Figures/total_time_sel.svg",width=12,height=6)

```
## Alt w stacked bars - separate for 7.5 hp courses

```{r}
plotdata <- enkat |>
  filter(!is.na(course_gr)) |>
#  mutate(subject = factor(subject, levels = c("Matte LTH", "Matte NF", 
#  "Matstat LTH", "Matstat NF"))) |>
  count(course_gr, Tid_totalt_timmar) |>
  group_by(course_gr) |>
  mutate(pct = n / sum(n), n_total = sum(n)) |>
  ungroup() |>
  mutate(course_gr_lbl = fct_inorder(paste0(course_gr, " (n = ", n_total, ")")))

plotdata |>
  ggplot(aes(x =course_gr_lbl, y = pct, fill = Tid_totalt_timmar)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_y_continuous(labels = percent) +
  labs(
    title="Total study hours",
    x = NULL,
    y = "Percent",
    fill = NULL
  ) +
  scale_fill_viridis_d(
    option = "D",
 #   labels = c("Inte alls", "","","", "", "Fullständigt")
  ) +
  scale_x_discrete(drop = FALSE) +
  guides(fill = guide_legend(nrow = 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
   # legend.justification = "center",
  #  legend.box.just = "top",
    legend.direction = "horizontal"
  )


```
(I prefer the standard bargraphs side by side)

## Scheduled time - as percentage of total course time



For all 
```{r}

nlbl= paste0(" n = ", enkat |>count()) 
enkat |>
  count(Tid_schema_grp5) |>
  mutate(pct = n / sum(n), n_total = sum(n)) |>
  ungroup() |>
#  mutate(nlbl = paste0(" (n = ", n_total, ")")) |>
  ggplot(aes(x = Tid_schema_grp5, y = pct, fill = Tid_schema_grp5)) +
  geom_col() +
  geom_text(aes(label = n), vjust = -0.4, size = 3) +
  scale_y_continuous(labels = scales::percent, expand = expansion(mult = c(0, .12))) +
  scale_fill_brewer(palette = "Greens", direction = 1) +
  labs(x = NULL, y = NULL,  
       title = "Proportion scheduled time (of own total in this course)",
       caption = str_wrap(attr(enkat$Tid_schema_proc, "label"),width=50),
       subtitle = nlbl)  +
  theme_minimal() +
  theme(legend.position = "none")
```

# Split by subject and faculty
```{r}
enkat |>
  filter(!is.na(subject)) |>
  mutate(subject = factor(subject, levels = c("Matte LTH", "Matte NF", "Matstat LTH", "Matstat NF"))) |>
  count(subject, Tid_schema_grp5) |>
  group_by(subject) |>
  mutate(pct = n / sum(n), n_total = sum(n)) |>
  ungroup() |>
  mutate(subject_lbl = fct_inorder(paste0(subject, " (n = ", n_total, ")"))) |>
  ggplot(aes(x = Tid_schema_grp5, y = pct, fill = Tid_schema_grp5)) +
  geom_col() +
  geom_text(aes(label = n), vjust = -0.4, size = 3) +
  scale_y_continuous(labels = scales::percent, expand = expansion(mult = c(0, .12))) +
  scale_fill_brewer(palette = "Greens", direction = 1) +
  facet_wrap(~ subject_lbl, nrow = 2) +
  labs(x = NULL, y = NULL, 
  title = "Proportion scheduled time (of own total in this course)",
       caption = str_wrap("Approximately, what is the percentage 
       of the total study time for this course 
       that you have spent on scheduled activities the last week? 
       Give an approximate answer in percentage. 
       The rest of the time is assumed to be self study.",
                          width=120)) +
  theme_minimal() +  
  theme(legend.position = "none")
```
## Separate for 7.5 hp courses (50% pace) B2, Matstat etc

```{r}
enkat |>
  filter(!is.na(course_gr)) |>
#  mutate(subject = factor(subject, levels = c("Matte LTH", "Matte NF", 
#  "Matstat LTH", "Matstat NF"))) |>
  count(course_gr, Tid_schema_grp5) |>
  group_by(course_gr) |>
  mutate(pct = n / sum(n), n_total = sum(n)) |>
  ungroup() |>
  mutate(course_gr_lbl = fct_inorder(paste0(course_gr, " (n = ", n_total, ")"))) |>
  ggplot(aes(x = Tid_schema_grp5, y = pct, fill = Tid_schema_grp5)) +
  geom_col() +
  geom_text(aes(label = n), vjust = -0.4, size = 3) +
  scale_y_continuous(labels = scales::percent, 
                     expand = expansion(mult = c(0, .12))) +
  scale_fill_brewer(palette = "Greens", direction = 1) +
  facet_wrap(~ course_gr_lbl, nrow = 2) +
  labs(x = NULL, y = NULL, 
       title = "Proportion scheduled time (of own total in this course)",
       caption = str_wrap("Approximately, what is the percentage 
       of the total study time for this course 
       that you have spent on scheduled activities the last week? 
       Give an approximate answer in percentage. 
       The rest of the time is assumed to be self study.",
                          width=120)) +
  theme_minimal() +
  theme(legend.position = "none")

```

## The same but as violins

All courses 
```{r}
#| label: violin_all

seldata =  enkat |> filter(!is.na(subject)) 
plotdata= bind_rows(
  seldata |> mutate(subject = "Total"),
  seldata
) |>
  mutate(subject = factor(subject,
    levels = c("Total", "Matte LTH", "Matte NF", "Matstat LTH", "Matstat NF")))

plotdata_sum <- plotdata |> group_by(subject) |> 
  summarize(n=n(),nlab=paste0("n = \n",n),Tid_schema_proc = mean(Tid_schema_proc))

nlbl=paste0(" n = ", plotdata |> filter(subject != "Total") |> count()) 

plotdata |>
  ggplot(aes(x = subject, y = Tid_schema_proc, fill = subject)) +
  geom_violin() +
 geom_text(aes(label = nlab), data = plotdata_sum) +
  scale_fill_manual(values = c("Total" = "grey70",
    "Matte LTH"   = "#74c476",
    "Matte NF"    = "#41ab5d",
    "Matstat LTH" = "#238b45",
    "Matstat NF"  = "#005a32")) +
  labs(x = NULL, y = NULL, title = "Proportion time on scheduled activities (%)", subtitle=nlbl,
    caption = str_wrap("Approximately, what is the percentage 
       of the total study time for this course 
       that you have spent on scheduled activities the last week? 
       Give an approximate answer in percentage. 
       The rest of the time is assumed to be self study.",
                          width=120)) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none",plot.caption = element_text(size=10))

```


### 7.5 hp courses ( = pace 50 %)


```{r}
#| label: violin_sel

seldata =  enkat |> filter(!is.na(course_gr)) 

plotdata= bind_rows(
  seldata |> mutate(course_gr = "Total"),
  seldata
) |>
  mutate(course_gr = factor(course_gr,
    levels = c("Total", levels(seldata$course_gr))))

plotdata_sum <- plotdata |> group_by(course_gr) |> 
  summarize(n=n(),nlab=paste0("n = \n",n),Tid_schema_proc = mean(Tid_schema_proc))

nlbl=paste0(" n = ", plotdata |> filter(course_gr != "Total") |> count()) 

plotdata |>
  ggplot(aes(x = course_gr, y = Tid_schema_proc, fill = course_gr)) +
  geom_violin() +
 geom_text(aes(label = nlab), data = plotdata_sum) +
  # scale_fill_manual(values = c("Total" = "grey70",
  #   "Matte LTH"   = "#74c476",
  #   "Matte NF"    = "#41ab5d",
  #   "Matstat LTH" = "#238b45",
  #   "Matstat NF"  = "#005a32")) +
   scale_fill_brewer(palette = "Greens", direction = 1) +
  labs(x = NULL, y = NULL, title = "Andel Schemalagd tid (%)", subtitle=nlbl,
     caption = str_wrap("Approximately, what is the percentage 
       of the total study time for this course 
       that you have spent on scheduled activities the last week? 
       Give an approximate answer in percentage. 
       The rest of the time is assumed to be self study.",
                          width=120)) +
  theme_minimal(base_size=14) +
  theme(legend.position = "none")
ggsave("Figures/violin_sel.emf",width=12,height=6)
ggsave("Figures/violin_sel.svg",width=12,height=6)

```
## AI usage

Results from VAR05_12 "Generativ AI (t ex Copilot, ChatGPT)"

### All answers
```{r}
nlbl= paste0(" n = ", enkat |>count()) 
cap=item_labels |> filter(var=="Resurs_Likert_12") |> pull(question)
enkat |>
  count(Resurs_Likert_12) |>
  mutate(pct = n / sum(n), n_total = sum(n)) |>
  ungroup() |>
#  mutate(nlbl = paste0(" (n = ", n_total, ")")) |>
  ggplot(aes(x = Resurs_Likert_12, y = pct, fill = Resurs_Likert_12)) +
  geom_col() +
  #geom_text(aes(label = n), vjust = -0.4, size = 3) +
  scale_x_discrete( labels = c("Never", "","","", "", "Always"))+
  scale_y_continuous(labels = scales::percent, expand = expansion(mult = c(0, .12))) +
  scale_fill_brewer(palette = "Greens", direction = 1) +
  labs(x = NULL, y = NULL,  title = "AI use",
       caption = str_wrap(cap,width=100),
       subtitle = nlbl)  +
  theme_minimal(base_size=15) +
  theme(legend.position = "none",
        plot.caption=element_text(size=10))

```


Per subject
```{r}
nlbl= paste0("n = ", enkat|> 
               filter(!is.na(subject),!is.na(Resurs_Likert_12)) |>
               count()) 
cap=item_labels |> filter(var=="Resurs_Likert_12") |> pull(question)
enkat  |> select(Resurs_Likert_12,subject) |>
  filter(!is.na(subject),!is.na(Resurs_Likert_12)) |>
  mutate(subject = factor(subject, levels = c("Matte LTH", "Matte NF", "Matstat LTH", "Matstat NF"))) |>
  count(subject, Resurs_Likert_12) |>
  group_by(subject) |> 
  mutate(pct=n/sum(n),n_gr_total=sum(n)) |> 
#  mutate(n_gr_total=sum(n)) |> 
 # ungroup() |> 
  #mutate(pct = n / n_gr_total) |>
#  ungroup() |>
  mutate(subject_lbl = fct_inorder(paste0(subject, " (n = ", n_gr_total, ")"))) |>
#  mutate(nlbl = paste0(" (n = ", n_total, ")")) |>
  ggplot(aes(x = Resurs_Likert_12, y = pct, fill = Resurs_Likert_12)) +
  geom_col(position=position_stack()) +
  #geom_text(aes(label = n), vjust = -0.4, size = 3) + 
  scale_x_discrete( labels = c("Never", "","","", "", "Always")) +
  scale_y_continuous(labels = scales::percent, expand = expansion(mult = c(0, .12))) +
  scale_fill_brewer(palette = "Greens", direction = 1) +
  facet_wrap(~ subject_lbl, nrow = 2) +
  labs(x = NULL, y = NULL,  title = "AI use",
       caption = str_wrap(cap,width=100),
       subtitle = nlbl)  +
  theme_minimal(base_size=15) +
  theme(legend.position = "none",
        plot.caption=element_text(size=10))



ggsave("Figures/AI_use_subject.emf",width=12,height=6)
ggsave("Figures/AI_use_subject.svg",width=12,height=6)
```

```{r}

enkat_sel <- enkat |> filter(!is.na(subject),!is.na(Resurs_Likert_1),
                             !is.na(Resurs_Likert_12),!is.na(Resurs_Likert_18))

plotdata_1 <- enkat_sel  |> select(Resurs_Likert_1,subject) |>
  mutate(subject = factor(subject, 
                          levels = c("Matte LTH", "Matte NF", 
                                     "Matstat LTH", "Matstat NF"))) |>
  count(subject, Resurs_Likert_1) |>
  group_by(subject) |> 
  mutate(pct=n/sum(n),n_gr_total=sum(n), var='Course Lit') |> 
#  mutate(subject_lbl = fct_inorder(paste0(subject, " (n = ", n_gr_total, ")"))) |> 
  rename(level=Resurs_Likert_1)

plotdata_12 <- enkat_sel  |> select(Resurs_Likert_12,Resurs_Likert_1,subject) |>
  mutate(subject = factor(subject, levels = c("Matte LTH", "Matte NF", "Matstat LTH", "Matstat NF"))) |>
  count(subject, Resurs_Likert_12) |>
  group_by(subject) |> 
  mutate(pct=n/sum(n),n_gr_total=sum(n), var='AI use') |> 
#  mutate(subject_lbl = fct_inorder(paste0(subject, " (n = ", n_gr_total, ")"))) |>  
  rename(level=Resurs_Likert_12)

plotdata_18 <- enkat_sel  |> select(Resurs_Likert_18,Resurs_Likert_1,subject) |>
  mutate(subject = factor(subject, levels = c("Matte LTH", "Matte NF", "Matstat LTH", "Matstat NF"))) |>
  count(subject, Resurs_Likert_18) |>
  group_by(subject) |> 
  mutate(pct=n/sum(n),n_gr_total=sum(n), var='tekniskfysik.org') |> 
#  mutate(subject_lbl = fct_inorder(paste0(subject, " (n = ", n_gr_total, ")"))) |> 
  rename(level=Resurs_Likert_18)


plotdata <- plotdata_1 |> bind_rows(plotdata_12) |> bind_rows(plotdata_18) |> 
    mutate(subject_lbl = fct_inorder(paste0(subject, "\n (n = ", n_gr_total, ")"))) 
#  mutate(nlbl = paste0(" (n = ", n_total, ")")) |>

plotdata |> 
  ggplot(aes(x = var, y = pct, fill = level )) +
  geom_col(position=position_stack()) +
  geom_hline(yintercept = 0.5) +
  #geom_text(aes(label = n), vjust = -0.4, size = 3) + 
 # scale_x_discrete( labels = c("Never", "","","", "", "Always")) +
  scale_y_continuous(labels = scales::percent, expand = expansion(mult = c(0, .12))) +
  scale_fill_brewer(palette = "Greens", direction = 1,labels = c("Never", "","","", "", "Always")) +
  facet_wrap(~ subject_lbl, nrow = 1) +
  labs(x = NULL, y = NULL,
       title = 'Use of resources for self-study',
       caption = str_wrap(cap,width=100),
     #  subtitle = nlbl,
       fill="How often\n last week?")  +
  theme_minimal(base_size=15) +
  theme(#legend.position = "none",
    axis.text.x=element_text(angle=45,hjust=1),
        plot.caption=element_text(size=10))


ggsave("Figures/AICLTF_use_subject.emf",width=12,height=6)
ggsave("Figures/AICLTF_use_subject.svg",width=12,height=6)
```

## Course literature
Per subject
```{r}
nlbl= paste0("n = ", enkat|> 
               filter(!is.na(subject),!is.na(Resurs_Likert_1)) |>
               count()) 
cap=item_labels |> filter(var=="Resurs_Likert_1") |> pull(question)
enkat |> select(Resurs_Likert_1,subject) |> 
  filter(!is.na(subject),!is.na(Resurs_Likert_1)) |>
  mutate(subject = factor(subject, levels = c("Matte LTH", "Matte NF", "Matstat LTH", "Matstat NF"))) |>
  count(subject, Resurs_Likert_1) |>
  group_by(subject) |> 
  mutate(pct=n/sum(n),n_gr_total=sum(n)) |> 
#  mutate(n_gr_total=sum(n)) |> 
 # ungroup() |> 
  #mutate(pct = n / n_gr_total) |>
#  ungroup() |>
  mutate(subject_lbl = fct_inorder(paste0(subject, " (n = ", n_gr_total, ")"))) |>
#  mutate(nlbl = paste0(" (n = ", n_total, ")")) |>
  ggplot(aes(x = Resurs_Likert_1, y = pct, fill = Resurs_Likert_1)) +
  geom_col() +
  #geom_text(aes(label = n), vjust = -0.4, size = 3) + 
  scale_x_discrete( labels = c("Never", "","","", "", "Always")) +
  scale_y_continuous(labels = scales::percent, expand = expansion(mult = c(0, .12))) +
  scale_fill_brewer(palette = "Greens", direction = 1) +
  facet_wrap(~ subject_lbl, nrow = 2) +
  labs(x = NULL, y = NULL,  title = "Course Literature",
       caption = str_wrap(cap,width=100),
       subtitle = nlbl)  +
  theme_minimal(base_size=15) +
  theme(legend.position = "none",
        plot.caption=element_text(size=10))



ggsave("Figures/CL_use_subject.emf",width=12,height=6)
ggsave("Figures/CL_use_subject.svg",width=12,height=6)
```


## Goals for studying mathematics

```{r}

past_lbl=item_labels |> select(var,item) |> 
filter(str_starts(var,"Påstående")) |> 
    filter(parse_number(var)!=13) |> 
  mutate(
    item = item |>
      str_remove("^Jag studerar för att\\s*") |>
      str_to_sentence(locale = "sv")
  ) |> pull(item)



plotdata= enkat |> 
  select(ID,prog,fakultet,course_code,subject,course_gr,starts_with("Påstående")) |> 
  pivot_longer(starts_with("Påstående"),names_to="Statement")  |> 
  filter(parse_number(Statement)!=13,!is.na(value)) |> 
  mutate(Statement_f = 
    factor(parse_number(Statement), ordered = TRUE,
                            labels = str_wrap(past_lbl,width=25))) |> 
  count(Statement_f, value, .drop = FALSE) |>
  group_by(Statement_f) |> 
  mutate(percent = n / sum(n)) |> 
      mutate(percent_full = 
              # max(percent * as.numeric(value=="Fullständigt"),na.rm=TRUE)) 
                max(percent * as.numeric(value=="Completely"),na.rm=TRUE)) 


plotdata |> 
  ggplot(aes(x = fct_reorder(Statement_f,-percent_full), 
             y = percent, 
             fill = value)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_y_continuous(labels = percent) +
  labs(
    title="Jag studerar för att...",
    x = NULL,
    y = "Procent",
    fill = NULL
  ) +
  scale_fill_viridis_d(
    option = "D",
    labels = c("Inte alls", "","","", "", "Fullständigt")
  ) +
  scale_x_discrete(drop = FALSE) +
  guides(fill = guide_legend(reverse = TRUE, nrow = 1)) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
   # legend.justification = "center",
  #  legend.box.just = "top",
    legend.direction = "horizontal"
  )


ggsave("Figures/goals.emf",width=12,height=6)
ggsave("Figures/goals.svg",width=12,height=6)

```

## Attendance at scheduled teaching
### Lectures
Attendance at lectures last week 
```{r}
nlbl=paste0("n = ",enkat |> filter(!is.na(course_gr)) |> count() |> pull(n)) 

  enkat |> filter(!is.na(course_gr)) |> 
  ggplot(aes(x=Tid_uppd_timmar_1)) + 
  geom_histogram( breaks = c(-0.5,0.5,2.5,4.5,6.5,8.5,10.5,12.5),
                  fill = 'darkgreen', alpha = 0.7) + 
    facet_wrap(~course_gr,scales='free_y')+
  labs(title = 'Attended Lectures',
       subtitle = paste0('7.5 hp courses, ',nlbl),

             x = "Time (hours)") +
  theme_minimal(base_size=15) 
                

```

## Resurser  - Andel på nivå 5 och 6
```{r}
# 
# enkat |> count(course_gr,Resurs_Likert_1_bin) |>
#   filter(!is.na(Resurs_Likert_1_bin)) |> 
#   group_by(course_gr) |>
#   mutate(total=sum(n),pct=n/total)

item_short <- c(
  Resurs_Likert_1  = "Course literature",
  Resurs_Likert_2  = "Videos LU teachers",
  Resurs_Likert_3  = "Videos other uni",
  Resurs_Likert_4  = "Videos non-uni",
  Resurs_Likert_5  = "Study friends",
  Resurs_Likert_6  = "Family/friends",
  Resurs_Likert_7  = "Lecture notes",
  Resurs_Likert_8  = "Old exams",
  Resurs_Likert_9  = "Discussion forum",
  Resurs_Likert_10 = "Other Canvas material",
  Resurs_Likert_11 = "Google/Wikipedia",
  Resurs_Likert_12 = "Generative AI",
  Resurs_Likert_13 = "Wolfram/GeoGebra/Desmos",
  Resurs_Likert_14 = "Python/R/Matlab",
  Resurs_Likert_15 = "Mozquizto",
  Resurs_Likert_16 = "Own lecture notes",
  Resurs_Likert_17 = "Other student's notes",
  Resurs_Likert_18 = "tekniskfysik.org",
  Resurs_Likert_19 = "Library",
  Resurs_Likert_20 = "Frågeforum (MC)",
  Resurs_Likert_21 = "Other online forum",
  Resurs_Likert_22 = "Email to teacher",
  Resurs_Likert_23 = "Other resource 1",
  Resurs_Likert_24 = "Other resource 2"
)


course_n <- enkat |> 
  filter(!is.na(course_gr)) |>
  count(course_gr) |>
  deframe()


plotdata <- enkat |>
  pivot_longer(
    matches("^Resurs_Likert_\\d+_bin$"),
    names_to  = "item",
    values_to = "bin"
  ) |>
  filter(item %in% paste0("Resurs_Likert_", 
                          c(1, 2, 5, 8, 12, 16, 18),
                          "_bin")) |>
  filter(!is.na(bin), !is.na(course_gr)) |>
  count(course_gr, item, bin) |>
  group_by(course_gr, item) |>
  mutate(total = sum(n), pct = n / total) |>
  ungroup() |>
  mutate(course_gr = fct_relabel(course_gr, \(x) paste0(x, " (n=", course_n[x], ")"))) |> 
  filter(bin=="Ofta") |> 
  mutate(item_label = item_short[
    str_remove(item, "_bin$")
    ]) 
  
  overall <- plotdata |>
    summarise(n = sum(n), total = sum(total), .by = item_label) |>
    mutate(pct = n / total,
           course_gr = factor("Overall", levels = c(levels(plotdata$course_gr), "Overall")))

  plotdata_full <- bind_rows(
    plotdata |> mutate(course_gr = fct_expand(course_gr, "Overall")),
    overall
  )

  plotdata_full |>
    filter(course_gr != "Overall") |>
    arrange(course_gr) |>
    ggplot(aes(x=pct, y=course_gr, color=item_label, group=item_label)) +
    geom_path() +
    geom_jitter(size=3, width=0.02, height=0, alpha=0.8) +
    geom_point(
      data = filter(plotdata_full, course_gr == "Overall"),
      shape = 18, size = 4
    ) +
    labs(title='Common resources', x = "Percentage", y = "Course", color = NULL) +
    scale_x_continuous(labels = scales::percent, limits = c(0, 1)) +
    scale_y_discrete(limits = rev, drop = FALSE) +
    theme_minimal(base_size = 15)
  
  

ggsave("Figures/resources_high.emf",width=9,height=6)
ggsave("Figures/resources_high.svg",width=9,height=6)
```

