---
title: "Resources for learning mathematics - results and figures for the publication"
author: "Linda Hartman and Sara Maad Sasane"
format: 
  html:
    embed-resources: true
echo: false
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(skimr)
library(scales)
library(patchwork)
library(gt)
```

## Data


```{r}
#| label: data

enkat_eng <- readRDS('Data/enkat_eng.rds') |> select(-starts_with("VAR"))
item_labels <- readRDS('Data/item_labels.rds')
ct <- readRDS('data/course_table.rds')
```

We had `r  dim(enkat_eng)[1]` respondents to the survey. Data was exported to SPSS and cleaned in script `dm.R`. 
One student had missing course-code, and is excluded from all analysis separated for course or subject, resulting in `r  dim(enkat_eng |> filter(!is.na(course_code)))[1]` students for these analyses.


For all discussion, unless distinctly stated, reference to *mathematics* courses includes both *mathematics* and *mathematical statistics*, unless distinctly stated otherwise (such as in graphs where we split *mathematics* from *mathematical statistics*).

Free text answers are not included in this report.




```{r}
#| label: skim
#enkat |>  skim()

```
## Table of courses/subject and faculty
This table gives an overview over the different courses, how they are distributed on faculty and subject, and the speed (% of full time) at which they are given.

```{r}
#| label: course_table

ct |> 
  select(course_code, course_name_abbr, faculty_abbr, subject,year,speed,comments) |> 
  arrange(faculty_abbr,desc(subject)) |> 
  gt()
```

For analysis of most common resources we include all answers, but for analysis on hours of time spent on the course in total, and on scheduled events we for comparability include only courses given at 50% of full speed.

## Goals for studying mathematics

```{r}
#| label: goals_bar
#| caption: Goals for studying this mathematic
past_lbl=item_labels |> select(var,item_eng) |> 
filter(str_starts(var,"Påstående")) |> 
    filter(parse_number(var)!=13) |> 
  mutate(
    item_eng = item_eng |>
      str_remove("^I study \\s*") |>
  #    str_remove("^I study because \\s*") |>
      str_to_sentence(locale = "en")
  ) |> pull(item_eng)

past_lbl[1]  <- "Enjoy mathematics"
past_lbl[2]  <- "Get a high grade"
past_lbl[3]  <- "Pass the course"
past_lbl[4]  <- "Understand the theory"
past_lbl[5]  <- "Improve computational skills"
past_lbl[6]  <- "Improve applied math skills"
past_lbl[7]  <- "Improve analytical skills"
past_lbl[8]  <- "Develop mathematical thinking"
past_lbl[9]  <- "Satisfy intellectual curiosity"
past_lbl[10] <- "Open my mind"
past_lbl[11] <- "Prepare for future work"
past_lbl[12] <- "Obtain my degree"

plotdata= enkat_eng |> 
  select(ID,prog,faculty,course_code,subject,course_gr,starts_with("Påstående")) |> 
  pivot_longer(starts_with("Påstående"),names_to="Statement")  |> 
  filter(parse_number(Statement)!=13,!is.na(value)) |> 
  mutate(Statement_f = 
    factor(parse_number(Statement), ordered = TRUE,
                            labels = past_lbl)) |> 
  count(Statement_f, value, .drop = FALSE) |>
  group_by(Statement_f) |> 
  mutate(percent = n / sum(n)) |> 
      mutate(percent_full = 
              # max(percent * as.numeric(value=="Fullständigt"),na.rm=TRUE)) 
                max(percent * as.numeric(value=="Completely"),na.rm=TRUE)) 


plotdata |> 
  ggplot(aes(x = fct_reorder(Statement_f,-percent_full), 
             y = percent, 
             fill = fct_rev(value))) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_y_continuous(labels = percent) +
  labs(
 #   title="I study to/because...",
     title="I study to...",
    caption="To what degree do you agree with the following statements about your studies in this course?",
    x = NULL,
   # y = "Percent",
   y=NULL,
    fill = NULL
  ) +
  scale_fill_viridis_d(
    option = "D",
    direction = -1,
    labels = c("Completely", "","","", "", "Not at all")
  ) +
  scale_x_discrete(drop = FALSE) +
  guides(fill = guide_legend(reverse = FALSE, nrow = 1)) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1,size=14),
    legend.position = "bottom",
   # legend.justification = "center",
  #  legend.box.just = "top",
    legend.direction = "horizontal",
    plot.margin = margin(5.5, 5.5, 5.5, 30)
  )


ggsave("Figures/article/goals.emf",width=13,height=6)
ggsave("Figures/article/goals.svg",width=13,height=6)

```

## Total hours spent on the course

The distribution of total hours spent in the course is presented for all students taking a (mathematics) course at speed = 50%:
```{r}

nlbl_all= paste0(" n = ", enkat_eng |> filter(!is.na(course_code),speed ==0.5) |> count()) 
p_total_all <- enkat_eng |>
  filter(!is.na(course_code)) |> 
  count(Tid_totalt_timmar) |>
  mutate(pct = n / sum(n), n_total = sum(n)) |>
  ungroup() |>
#  mutate(nlbl = paste0(" (n = ", n_total, ")")) |>
  ggplot(aes(x = Tid_totalt_timmar, y = pct, fill = Tid_totalt_timmar)) +
  geom_col(color="black") +
#  geom_text(aes(label = n), vjust = -0.4, size = 3) +
  scale_y_continuous(labels = scales::percent, expand = expansion(mult = c(0, .12))) +
  scale_fill_brewer(palette = "Greens", direction = 1) +
  labs(x = 'Hours', y = 'Percentage',  title = "Total hours spent in this course",
       caption = str_wrap("Approximately, how many hours have you spent on your
                          mathematics studies in this course totally the last week?\n 
                          The question refers to the total time spent, 
                          consisting of scheduled time as well as 
                          self study.",width=110),
       subtitle = paste0('Courses at 50% speed, ',nlbl_all))  +
       #subtitle = paste0('Scheduled teaching + self studies, ',nlbl_all))  +
  theme_minimal(base_size=14) +
  theme(legend.position = "none")
p_total_all


ggsave("Figures/article/studytime.emf",width=13,height=6)
ggsave("Figures/article/studytime.svg",width=13,height=6)
```



## Results, time on different activities

### Time allocated to scheduled activities (%)
The stated proportion of time allocated to scheduled activities is reported for all students with data on which course they took (1 student lacking that info).
```{r}
#| label: violin_all

seldata =  enkat_eng |> filter(!is.na(subj_fac)) 
plotdata= bind_rows(
  seldata |> mutate(subj_fac = "Total"),
  seldata) |>
  mutate(subj_fac = factor(subj_fac,
                           levels = c("Total", "Math LTH",  "Math Sci", "MStat LTH", "MStat Sci")))



plotdata_sum <- plotdata |> group_by(subj_fac) |> 
  summarize(n=n(),nlab=paste0("n = \n",n),Tid_schema_proc = mean(Tid_schema_proc))

nlbl=paste0(" n = ", plotdata |> filter(subj_fac != "Total") |> count()) 

plotdata |>
  ggplot(aes(x = subj_fac, y = Tid_schema_proc, fill = subj_fac)) +
  geom_violin() +
  geom_text(aes(label = nlab, color = subj_fac), data = plotdata_sum) +
  scale_color_manual(values = 
                       c("Total" = "white",
                         "Math LTH"   = "black",
                         "Math Sci"    = "black",
                         "MStat LTH" = "black",
                         "MStat Sci"  = "black"), guide = "none") +
  scale_fill_manual(values = 
                      c("Total"     = "#005a32",
                        "Math LTH"   = "#238b45",
                        "Math Sci"    = "#41ab5d",
                        "MStat LTH" = "#74c476",
                        "MStat Sci"  = "#bae4b3")) +
  labs(x = NULL, y = NULL, title = "Proportion time student spent on scheduled activities (%)", subtitle=nlbl,
       caption = str_wrap("Approximately, what is the percentage 
       of the total study time for this course 
       that you have spent on scheduled activities the last week? 
       Give an approximate answer in percentage. 
       The rest of the time is assumed to be self study.",
                          width=120)) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none",plot.caption = element_text(size=10))

```

### Time allocated to study (hours) during last week on course activities


Of the `r enkat_eng |> filter(!is.na(subj_fac), speed==0.5) |>  count()` students taking a mathematics course at 50% speed, 
`r enkat_eng |> filter(!is.na(subj_fac), speed==0.5) |> filter_out(!is.na(Tid_uppd_timmar_4), !is.na(Tid_uppd_timmar_5), !is.na(Tid_uppd_timmar_6)) |>  count()`
 did not provide data on hours spent on different scheduled activities. Thus the following graphs represent 
`r enkat_eng |> filter(!is.na(subj_fac), speed==0.5, !is.na(Tid_uppd_timmar_1), !is.na(Tid_uppd_timmar_2), !is.na(Tid_uppd_timmar_3),  !is.na(Tid_uppd_timmar_4), !is.na(Tid_uppd_timmar_5), !is.na(Tid_uppd_timmar_6)) |>  count()` 
students.
           
The question(s) that the students answered was:           
*For each of the activities below, please tell us approximately how many hours during the last week you spent on each of the listed activities in this course on average per week.* 

The question had six sub-questions, one for each activity (Lectures, Seminars, Exercise classes, Supplemental instruction, Computer exercises, Other teacher-led activity).

As the meaning of *exercise classes*, *seminars* and *computer exercises* differ between courses we report the sum of them together.

We did not include hours of supplemental instruction in the calculations, as that is run by older students without common planning with the course teachers, and is not considered teaching, but rather as support between students.

### Lectures, Exercises, and Combined Hours
Below we report the number of hours for Lectures, Exercise+Seminar (that includes computer exercises and other teacher-led activities), and the sum of both (= all teacher-led activities).



```{r}
#| label: att_lect_exc
#| message: false
#| warning: false
#| fig-height: 10
#| fig-width: 8

sel_data <- enkat_eng |>
  filter(!is.na(subj_fac), 
         speed==0.5,
         !is.na(Tid_uppd_timmar_1), !is.na(Tid_uppd_timmar_2), !is.na(Tid_uppd_timmar_3),
         !is.na(Tid_uppd_timmar_4), !is.na(Tid_uppd_timmar_5), !is.na(Tid_uppd_timmar_6),) |>
  mutate(Tid_exercise = Tid_uppd_timmar_2 + Tid_uppd_timmar_3 + Tid_uppd_timmar_5 + Tid_uppd_timmar_6)

nlbl <- paste0("n = ", sel_data |> count() |> pull(n))

# Add sample sizes for each group
group_ns <- bind_rows(
  tibble(group = "Total", n = nrow(sel_data)),
  sel_data |> group_by(subj_fac) |> count() |> 
    mutate(group = as.character(subj_fac)) |> select(group, n)
) |>
  mutate(
    group = factor(group, levels = c("Total", "Math LTH", "Math Sci", "MStat LTH", "MStat Sci")),
    group_label = paste0(group, "\n(n = ", n, ")")
  )

# Prepare data for lectures, exercises, and total
plot_data <- bind_rows(
  # Lectures
  sel_data |>
    mutate(hours = Tid_uppd_timmar_1, type = "Lectures", group = "Total") |>
    select(type, hours, group),
  # Exercises (seminars + exercise sessions + computer exercises)
  sel_data |>
    mutate(hours = Tid_exercise, type = "Exercises", group = "Total") |>
    select(type, hours, group),
  # Lect + Exercise combined
  sel_data |>
    mutate(hours = Tid_uppd_timmar_1 + Tid_exercise) |>
    mutate(type = "Lect + Exerc", group = "Total") |>
    select(type, hours, group),
  # By subject - lectures
  sel_data |>
    mutate(hours = Tid_uppd_timmar_1, type = "Lectures", group = as.character(subj_fac)) |>
    select(type, hours, group),
  # By subject - exercises
  sel_data |>
    mutate(hours = Tid_exercise, type = "Exercises", group = as.character(subj_fac)) |>
    select(type, hours, group),
  # By subject - combined
  sel_data |>
    mutate(hours = Tid_uppd_timmar_1 + Tid_exercise) |>
    mutate(type = "Lect + Exerc", group = as.character(subj_fac)) |>
    select(type, hours, group)
) |>
  mutate(
    group = factor(group, levels = c("Total", "Math LTH", "Math Sci", "MStat LTH", "MStat Sci")),
    type = factor(type, levels = c("Lectures", "Exercises", "Lect + Exerc"))
  ) |>
  left_join(group_ns |> select(group, group_label), by = "group") |>
  # Order group_label by group factor levels (which are already ordered)
  mutate(group_label = factor(group_label, levels = unique(group_ns$group_label)))

# Calculate excluded counts (observations beyond x-axis limit of 25)
excluded_counts <- bind_rows(
  # Lectures - Total
  sel_data |>
    mutate(hours = Tid_uppd_timmar_1, type = "Lectures", group = "Total") |>
    group_by(group, type) |>
    summarize(n_out = sum(hours > 25), .groups = "drop"),
  # Exercises - Total
  sel_data |>
    mutate(hours = Tid_exercise, type = "Exercises", group = "Total") |>
    group_by(group, type) |>
    summarize(n_out = sum(hours > 25), .groups = "drop"),
  # Lect + Exerc - Total
  sel_data |>
    mutate(hours = Tid_uppd_timmar_1 + Tid_exercise, type = "Lect + Exerc", group = "Total") |>
    group_by(group, type) |>
    summarize(n_out = sum(hours > 25), .groups = "drop"),
  # Lectures - By subject
  sel_data |>
    mutate(hours = Tid_uppd_timmar_1, type = "Lectures", group = as.character(subj_fac)) |>
    group_by(group, type) |>
    summarize(n_out = sum(hours > 25), .groups = "drop"),
  # Exercises - By subject
  sel_data |>
    mutate(hours = Tid_exercise, type = "Exercises", group = as.character(subj_fac)) |>
    group_by(group, type) |>
    summarize(n_out = sum(hours > 25), .groups = "drop"),
  # Lect + Exerc - By subject
  sel_data |>
    mutate(hours = Tid_uppd_timmar_1 + Tid_exercise, type = "Lect + Exerc", group = as.character(subj_fac)) |>
    group_by(group, type) |>
    summarize(n_out = sum(hours > 25), .groups = "drop")
) |>
  mutate(
    group = factor(group, levels = c("Total", "Math LTH", "Math Sci", "MStat LTH", "MStat Sci")),
    type = factor(type, levels = c("Lectures", "Exercises", "Lect + Exerc"))
  ) |>
  left_join(group_ns |> select(group, group_label), by = "group") |>
  # Order group_label by group factor levels
  mutate(group_label = factor(group_label, levels = unique(group_ns$group_label))) |>
  mutate(label = if_else(n_out > 0, paste0("n_out = ", n_out), NA_character_))

# Create plot with x-scale 0-20 for all columns
plot_data |>
  ggplot(aes(x = hours, fill = group, y = after_stat(100 * count / tapply(count, PANEL, sum)[PANEL]))) +
  geom_histogram(breaks = c(-0.5, 0.5, 2.5, 4.5, 6.5, 8.5, 10.5, 12.5, 14.5, 16.5, 18.5, 20.5, 22.5, 24.5),
                 color = 'black', linewidth = 0.3) +
  scale_x_continuous(limits = c(-1, 26)) +
  facet_grid(group_label ~ type, scales = "free_y") +
  scale_fill_manual(values = 
                      c("Total"     = "#005a32",
                        "Math LTH"   = "#238b45",
                        "Math Sci"    = "#41ab5d",
                        "MStat LTH" = "#74c476",
                        "MStat Sci"  = "#bae4b3"),
                    guide = "none") +
  geom_text(data = excluded_counts,
            aes(x = 19, y = Inf, label = label, fill = NULL),
            vjust = 1.2, hjust = 0.9, size = 3.5, fontface = "italic") +
  labs(title = 'Attendance at Lectures and Exercises',
       subtitle = paste0('Courses at speed = 50%, ', nlbl),
       x = "Time (hours)",
       y = "Percentage (%)") +
  theme_minimal(base_size = 13) +
  theme(panel.spacing = unit(0.5, "lines"),
        aspect.ratio = 1,
        strip.text = element_text(face = "bold"))

ggsave("Figures/article/hist_lect_exc.emf", width = 10, height = 12)
ggsave("Figures/article/hist_lect_exc.png", width = 10, height = 12)
```



## Use of self study resources

### Resources  - Andel på nivå 5 och 6


A graph of the most commonly used self study resources. The proportion of students stating level 5 or 6 for 7 most used resources, split by subject/faculty-combination.

```{r}
#| label: common_resources


item_short <- c(
  Resurs_Likert_1  = "Course literature",
  Resurs_Likert_2  = "Videos LU teachers",
  Resurs_Likert_3  = "Videos other uni",
  Resurs_Likert_4  = "Videos non-uni",
  Resurs_Likert_5  = "Study friends",
  Resurs_Likert_6  = "Family/friends",
  Resurs_Likert_7  = "Lecture notes",
  Resurs_Likert_8  = "Old exams",
  Resurs_Likert_9  = "Discussion forum",
  Resurs_Likert_10 = "Other Canvas material",
  Resurs_Likert_11 = "Google/Wikipedia",
  Resurs_Likert_12 = "Generative AI",
  Resurs_Likert_13 = "Wolfram/GeoGebra/Desmos",
  Resurs_Likert_14 = "Python/R/Matlab",
  Resurs_Likert_15 = "Mozquizto",
  Resurs_Likert_16 = "Own lecture notes",
  Resurs_Likert_17 = "Other student's notes",
  Resurs_Likert_18 = "tekniskfysik.org",
  Resurs_Likert_19 = "Library",
  Resurs_Likert_20 = "Frågeforum (MC)",
  Resurs_Likert_21 = "Other online forum",
  Resurs_Likert_22 = "Email to teacher",
  Resurs_Likert_23 = "Other resource 1",
  Resurs_Likert_24 = "Other resource 2"
)
```


```{r}
subj_fac_n <- enkat_eng |> 
  filter(!is.na(subj_fac)) |>
  count(subj_fac) |>
  deframe()


plotdata <- enkat_eng |>
  select(subj_fac,starts_with("Resurs_Likert")) |>
  filter(!is.na(subj_fac)) |>
  pivot_longer(
    matches("^Resurs_Likert_\\d+_bin$"),
    names_to  = "item",
    values_to = "bin"
  ) |> 
  filter(item %in% paste0("Resurs_Likert_", 
                          c(1, 2, 5, 8, 12, 16, 18),
                          "_bin")) |>
  filter(!is.na(bin), !is.na(subj_fac)) |>
  count(subj_fac, item, bin) |>
  # Set pct to 0 for missing combinations
   complete(subj_fac, item, bin, fill = list(n=0)) |> 
  group_by(subj_fac, item) |>
  mutate(total = sum(n), pct = n / total) |>
  ungroup() |>
  mutate(subj_fac = fct_relabel(subj_fac, \(x) paste0(x, " (n=", subj_fac_n[x], ")"))) |> 
  filter(bin=="Often") |> 
  mutate(item_label = item_short[
    str_remove(item, "_bin$")
    ]) 
 
  # Calculate total sample size for overall
  overall_n <- enkat_eng |> 
    filter(!is.na(subj_fac)) |> 
    nrow()
  
  overall <- plotdata |>
    summarise(n = sum(n), total = sum(total), .by = item_label) |>
    mutate(pct = n / total,
           subj_fac = factor(paste0("Overall (n=", overall_n, ")"), 
                            levels = c(levels(plotdata$subj_fac), 
                                       paste0("Overall (n=", overall_n, ")"))),
           item_label0 = fct_reorder(item_label, pct, .desc = TRUE)) |> 
    mutate(
    item_label = factor(paste0(item_label0, " (", round(pct * 100), "%)")),
    item_label = fct_reorder(item_label, pct, .desc = TRUE)) 
   
  
  
plotdata <- plotdata |>
  mutate(
    item_label = factor(item_label, levels = levels(overall$item_label0),labels=levels(overall$item_label))
  )

#  plotdata_full <- bind_rows(
#    plotdata |> mutate(subj_fac = fct_expand(subj_fac, "Overall")),
#    overall
#  )
  
```


```{r}
#| label: resources_5_6
#| fig-width: 9
#| fig-height: 6
#| message: false
#| warning: false

                                      
set.seed(123)
plotdata |>
    ggplot(aes(x=pct, y=subj_fac, color=item_label)) +
    geom_path(aes(y=fct_rev(subj_fac),group=item_label), alpha=0.7, linewidth=1.2) +
    geom_jitter(aes(y=fct_rev(subj_fac)), size=5, width=0.005, height=0, alpha=0.8) +
    geom_point(
      data = overall,
      aes(group=NA),
      shape = 18, size = 8
    ) +
    labs(title='Commonly used self study resources', 
         subtitle = 'Overall percentage for level 5 and 6 in parentheses',
         x = "Percentage for level 5+6\n (1 = Never, 6 = Always)", 
         y = NULL, 
         color = NULL,
         ) +
    scale_x_continuous(labels = scales::percent, limits = c(0, 1)) +
    scale_y_discrete(drop = FALSE) +
    theme_minimal(base_size = 15)
  
  

ggsave("Figures/article/resources_high.emf",width=9,height=6)
ggsave("Figures/article/resources_high.svg",width=9,height=6)
```

### AI use 

## Appendix
### Attendance at scheduled hours 
Alternatives with lectures and exercises separately (no total). Code (commented out) is available here in the qmd-file. 

#### Lectures
Attendance at lectures last week
```{r}
#| label: att_lectures
# 
# seldata =  enkat_eng |>
#   filter(!is.na(subj_fac),!is.na(Tid_uppd_timmar_1),course_code!='FMSF80')
# 
# nlbl=paste0("n = ",seldata |> count() |> pull(n))
# 
# seldata |>
#   add_count(subj_fac, name = "n_total") |>
#   arrange(subj_fac) |>
#   mutate(subj_fac_lbl = fct_inorder(paste0(subj_fac, " (n = ", n_total, ")"))) |>
#   ggplot(aes(x=Tid_uppd_timmar_1)) +
#   geom_histogram( breaks = c(-0.5,0.5,2.5,4.5,6.5,8.5,10.5,12.5),
#                   fill = 'darkgreen', alpha = 0.7,color='black') +
#   facet_wrap(~subj_fac_lbl,scales='free_y')+
#   labs(title = 'Attended Lectures',
#        subtitle = paste0('Course FMSF80 excluded, ',nlbl),
#        x = "Time (hours)") +
#   theme_minimal(base_size=15)
# 
# ggsave("Figures/article/hist_lect.emf",width=12,height=6)
# ggsave("Figures/article/hist_lect.svg",width=12,height=6)
```

#### Exercises
Attendance at exercises last week
```{r}
#| label: att_exercises
# 
# seldata <-   enkat_eng |>
#   filter(!is.na(subj_fac),!is.na(Tid_uppd_timmar_3),course_code!='FMSF80')
# 
# nlbl=paste0("n = ",seldata |> count() |> pull(n))
# 
# seldata |>
#   add_count(subj_fac, name = "n_total") |>
#   arrange(subj_fac) |>
#   mutate(subj_fac_lbl = fct_inorder(paste0(subj_fac, " (n = ", n_total, ")"))) |>
#   ggplot(aes(x=Tid_uppd_timmar_3)) +
#   geom_histogram( breaks = c(-0.5,0.5,2.5,4.5,6.5,8.5,10.5,12.5),
#                   fill = 'darkgreen', alpha = 0.7,color='black') +
#   facet_wrap(~subj_fac_lbl,scales='free_y')+
#   labs(title = 'Attended Exercise Sessions',
#        subtitle = paste0('Course FMSF80 excluded, ',nlbl),
#        x = "Time (hours)") +
#   theme_minimal(base_size=15)
# 
# ggsave("Figures/article/hist_exc.emf",width=12,height=6)
# ggsave("Figures/article/hist_exc.svg",width=12,height=6)
```


