---
title: "Resources for learning mathematics - results and figures for the publication"
author: "Linda Hartman and Sara Maad Sasane"
format: 
  html:
    embed-resources: true
echo: false
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(skimr)
library(scales)
library(patchwork)
library(gt)
```

## Data

Data from 585 respondents to a survey. Data was exported to SPSS and cleaned in script `dm.R`. 

Free text answers are not included here.


```{r}
#| label: data

enkat_eng <- readRDS('Data/enkat_eng.rds') |> select(-starts_with("VAR"))
item_labels <- readRDS('Data/item_labels.rds')
ct <- readRDS('data/course_table.rds')
```


```{r}
#| label: skim
#enkat |>  skim()

```
## Table of courses/subject and faculty
This table gives an overview over the different courses, how they are distributed on faculty and subject
```{r}
#| label: course_table

ct |>  gt()
```


## Goals for studying mathematics

```{r}
#| label: goals_bar

past_lbl=item_labels |> select(var,item_eng) |> 
filter(str_starts(var,"Påstående")) |> 
    filter(parse_number(var)!=13) |> 
  mutate(
    item_eng = item_eng |>
      str_remove("^I study \\s*") |>
  #    str_remove("^I study because \\s*") |>
      str_to_sentence(locale = "en")
  ) |> pull(item_eng)

past_lbl[1]  <- "Enjoy mathematics"
past_lbl[2]  <- "Get a high grade"
past_lbl[3]  <- "Pass the course"
past_lbl[4]  <- "Understand the theory"
past_lbl[5]  <- "Improve computational skills"
past_lbl[6]  <- "Improve applied math skills"
past_lbl[7]  <- "Improve analytical skills"
past_lbl[8]  <- "Develop mathematical thinking"
past_lbl[9]  <- "Satisfy intellectual curiosity"
past_lbl[10] <- "Open my mind"
past_lbl[11] <- "Prepare for future work"
past_lbl[12] <- "Obtain my degree"

plotdata= enkat_eng |> 
  select(ID,prog,faculty,course_code,subject,course_gr,starts_with("Påstående")) |> 
  pivot_longer(starts_with("Påstående"),names_to="Statement")  |> 
  filter(parse_number(Statement)!=13,!is.na(value)) |> 
  mutate(Statement_f = 
    factor(parse_number(Statement), ordered = TRUE,
                            labels = past_lbl)) |> 
  count(Statement_f, value, .drop = FALSE) |>
  group_by(Statement_f) |> 
  mutate(percent = n / sum(n)) |> 
      mutate(percent_full = 
              # max(percent * as.numeric(value=="Fullständigt"),na.rm=TRUE)) 
                max(percent * as.numeric(value=="Completely"),na.rm=TRUE)) 


plotdata |> 
  ggplot(aes(x = fct_reorder(Statement_f,-percent_full), 
             y = percent, 
             fill = fct_rev(value))) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_y_continuous(labels = percent) +
  labs(
 #   title="I study to/because...",
     title="I study to...",
    caption="To what degree do you agree with the following\nstatements about your studies in this course?",
    x = NULL,
   # y = "Percent",
   y=NULL,
    fill = NULL
  ) +
  scale_fill_viridis_d(
    option = "D",
    direction = -1,
    labels = c("Completely", "","","", "", "Not at all")
  ) +
  scale_x_discrete(drop = FALSE) +
  guides(fill = guide_legend(reverse = FALSE, nrow = 1)) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1,size=14),
    legend.position = "bottom",
   # legend.justification = "center",
  #  legend.box.just = "top",
    legend.direction = "horizontal",
    plot.margin = margin(5.5, 5.5, 5.5, 30)
  )


ggsave("Figures/goals_eng.emf",width=13,height=6)
ggsave("Figures/goals_eng.svg",width=13,height=6)

```

## Total hours spent on the course

For all 
```{r}

nlbl_all= paste0(" n = ", enkat_eng |>count()) 
p_total_all <- enkat_eng |>
  count(Tid_totalt_timmar) |>
  mutate(pct = n / sum(n), n_total = sum(n)) |>
  ungroup() |>
#  mutate(nlbl = paste0(" (n = ", n_total, ")")) |>
  ggplot(aes(x = Tid_totalt_timmar, y = pct, fill = Tid_totalt_timmar)) +
  geom_col(color="black") +
#  geom_text(aes(label = n), vjust = -0.4, size = 3) +
  scale_y_continuous(labels = scales::percent, expand = expansion(mult = c(0, .12))) +
  scale_fill_brewer(palette = "Greens", direction = 1) +
  labs(x = 'Hours', y = 'Percentage',  title = "Total hours spent in this course",
       caption = str_wrap("Approximately, how many hours have you spent on your
                          mathematics studies in this course totally the last week?\n 
                          The question refers to the total time spent, 
                          consisting of scheduled time as well as 
                          self study.",width=110),
       subtitle = paste0('Scheduled teaching + self studies, ',nlbl_all))  +
  theme_minimal(base_size=14) +
  theme(legend.position = "none")
p_total_all
```

### Proportion scheduled time - as Violins

#### Split by Subject & faculty
```{r}
#| label: violin_all

seldata =  enkat_eng |> filter(!is.na(subj_fac)) 
plotdata= bind_rows(
  seldata |> mutate(subj_fac = "Total"),
  seldata) |>
  mutate(subj_fac = factor(subj_fac,
    levels = c("Total", "Math, LTH",  "Math, Sci", "MStat, LTH", "MStat, Sci")))

    

plotdata_sum <- plotdata |> group_by(subj_fac) |> 
  summarize(n=n(),nlab=paste0("n = \n",n),Tid_schema_proc = mean(Tid_schema_proc))

nlbl=paste0(" n = ", plotdata |> filter(subj_fac != "Total") |> count()) 

plotdata |>
  ggplot(aes(x = subj_fac, y = Tid_schema_proc, fill = subj_fac)) +
  geom_violin() +
 geom_text(aes(label = nlab, color = subj_fac), data = plotdata_sum) +
  scale_color_manual(values = 
    c("Total" = "white",
    "Math, LTH"   = "black",
    "Math, Sci"    = "black",
    "MStat, LTH" = "black",
    "MStat, Sci"  = "black"), guide = "none") +
  scale_fill_manual(values = 
    c("Total"     = "#005a32",
    "Math, LTH"   = "#238b45",
    "Math, Sci"    = "#41ab5d",
    "MStat, LTH" = "#74c476",
    "MStat, Sci"  = "#bae4b3")) +
  labs(x = NULL, y = 'Proportion (%)', title = "Proportion time students spent on scheduled activities (%)", subtitle=nlbl,
    caption = str_wrap("Approximately, what is the percentage 
       of the total study time for this course 
       that you have spent on scheduled activities the last week? 
       Give an approximate answer in percentage. 
       The rest of the time is assumed to be self study.",
                          width=118)) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none",plot.caption = element_text(size=10))

```

